---
title: "Lesson 1 - Bio"
author: "Jeppe Dreyer Matzen"
date: "1/31/2021"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(scales)
library(haven) # Import Stata
```

# Exercise 1. T-test - Two independent group

```{r}

smok <- read_dta("Data/KS7_2_BirthW_Smoking.dta") %>% 
    janitor::clean_names() 
head(KS7_2_BirthW_Smoking) 
```


## a. The scale of the two variables?
```{r}
smok %>% 
    summary()
```


## b. Descriptive statistics: Draw by means of Stata a Box plot - what do you see from the plot?

## c. Test: Investigate by means of a hypothesis test in Stata if there is a significant difference in the means among smokers and non-smokers (α = 1%)

including

- what is the reason for using this method?

- assumptions

- hypotheses

- test statistic and distribution

- p-value

- what is the conclusion?

## d. Find a 99 % confidence interval for the difference between the means (see output from question c above)

- Interpret the confidence interval and compare it with the conclusion on the hypothesis test

## e. Assumptions: Is Weight normal distributed and are there variance homogeneity (in both groups)? 